# Interim version !NOT FINAL!

# Mini research

## Setting up the data

### Libraries

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```

### Data loading

#### UK 1974-2017

```{r}
accidents_data <- read.csv("Accidents.csv")
bikers_data <- read.csv("Bikers.csv")

missing_rows <- anti_join(bikers_data, accidents_data, by = "Accident_Index")

bikers_data <- filter(bikers_data, !Accident_Index %in% missing_rows$Accident_Index)

selected_col <- cbind(accidents_data[, c("Accident_Index",
                                     "Date",
                                     "Time",
                                     "Day",
                                     "Road_type",
                                     "Speed_limit",
                                     "Road_conditions",
                                     "Weather_conditions",
                                     "Light_conditions")],
                                      bikers_data[, "Severity"])

colnames(selected_col)[ncol(selected_col)] <- "Severity"
```

#### Austin 2010-2017

```{r}
my_data <- read.csv("Austin Bicycle Crashes 2010-2017.csv")


new_df <- my_data[, c(          "Crash.Time",
                                "Day.of.Week",
                                
                                "Light.Condition",
                                "Weather.Condition",
                                
                                "Speed.Limit",
                                "Average.Daily.Traffic.Amount",
                                "Manner.of.Collision",
                                "Traffic.Control.Type",
                                "Intersection.Related",
                                
                                "Person.Helmet",
                                "Crash.Severity")]

colnames(my_data)
```

#### Motorbikes

```{r}
data=read.csv("out.csv")
data = data[data$BikeAge!="999",]
data$BikeAge = as.numeric(as.character(data$BikeAge))
data = data[!is.na(data$BikeAge),]
data
```

### Cleaning the data

```{r}
remove_unknown_rows <- function(dataframe) {
  for (col in colnames(dataframe)) {
    dataframe <- dataframe[!grepl("Unknown", dataframe[[col]], ignore.case = TRUE), ]
    dataframe <- dataframe[!grepl("Missing data", dataframe[[col]], ignore.case = TRUE), ]
  }
  return(dataframe)
}
tidy_num <- remove_unknown_rows(selected_col)
```

### Selecting data

```{r}
selected_col_num <- cbind(tidy_num[, c("Accident_Index",
                                             "Date",
                                             "Time",
                                             "Day",
                                             "Speed_limit",
                                             "Light_conditions")])
```

## Converting to numeric

### Lists for numerisation

#### UK

```{r}
Road_type_list <- c("Slip road", # denoted by 1
                    "One way sreet", # denoted by 2
                    "Single carriageway", # denoted by 3
                    "Dual carriageway", # denoted by 4
                    "Roundabout") # denoted by 5

Road_conditions_list <- c("Dry", # denoted by 1
                          "Wet", # denoted by 2
                          "Snow", # denoted by 3
                          "Frost", # denoted by 4
                          "Flood") # denoted by 5

Weather_conditions_list <- c("Clear", # denoted by 1
                             "Clear and windy", # denoted by 2
                             "Fog", # denoted by 3
                             "Rain", # denoted by 4
                             "Rain and windy", # denoted by 5
                             "Snow", # denoted by 6
                             "Snow and windy", # denoted by 7
                             "Other") # denoted by 8

Light_conditions_list <- c("Daylight", # denoted by 1
                           "Darkness lights lit", # denoted by 2
                           "Darkness no lights") # denoted by 3

Severity_list <- c("Slight", # denoted by 1
                   "Serious", # denoted by 2
                   "Fatal") # denoted by 3
```

#### Austin

```{r}
#Let`s see what is inside Crash.Severity column
data.frame(table(new_df$Crash.Severity))

```

#### Motorbikes

```{r}
#tbd
```

### Numerisation

#### Function

```{r}
convert_to_numeric <- function(selected_col, column_name, value_list) {
  unique_values <- unique(selected_col[[column_name]])
  
  name_numeric <- paste0(column_name, "_num")
  cat("\nUnique values in ", column_name, ":\n")
  
  # Print custom mapping
  for (i in seq_along(value_list)) {
    cat(i, " = ", value_list[i], "\n")
  }
  
  selected_col <- selected_col %>%
    mutate(!!name_numeric := match(selected_col[[column_name]], value_list))
  
  return(selected_col[, name_numeric, drop = FALSE])
}
```

#### UK

```{r}
cat("Translations: \n")
selected_col_num <- cbind(selected_col_num, convert_to_numeric(tidy_num, 
                                                               "Road_conditions",
                                                               Road_conditions_list))
selected_col_num <- cbind(selected_col_num, convert_to_numeric(tidy_num, 
                                                               "Weather_conditions",
                                                               Weather_conditions_list))
selected_col_num <- cbind(selected_col_num, convert_to_numeric(tidy_num, 
                                                               "Light_conditions",
                                                               Light_conditions_list))
selected_col_num <- cbind(selected_col_num, convert_to_numeric(tidy_num, 
                                                               "Severity",
                                                               Severity_list))
```

#### Austin

```{r}
# Assign each injury a numeric value
new_df$Crash.Numeric <- ifelse(new_df$Crash.Severity=="Not Injured", 1,
                    ifelse(new_df$Crash.Severity=="Possible Injury", 2,
                    ifelse(new_df$Crash.Severity=="Non-Incapacitating Injury", 3, 
                    ifelse(new_df$Crash.Severity=="Incapacitating Injury", 4,
                    ifelse(new_df$Crash.Severity=="Killed", 5, 6)))))
# Delete unknown values
new_df <- subset(new_df, Crash.Numeric != 6)
```

#### Motorbikes

```{r}

```

## Stats

### UK

#### Daily crash frequency

```{r}
data_UK <- data.frame(table(selected_col_num$Day))
data_UK <- data_UK[order(-data_UK$Freq), ]

# Increase the bottom and right margins
par(mar=c(6, 4, 3, 3))  # Adjust the second and fourth values as needed

barplot(data_UK$Freq, names.arg = data_UK$Var1, 
        las = 2, col = "skyblue", 
        main = "Accident Counts by Day of the week", 
        xlab = "", 
        ylab = "",
        ylim = c(0, 150000))

# Add the number of accidents above each column with adjusted x-coordinates
text(x = seq_along(data_UK$Var1)*1.22-0.55, y = data_UK$Freq + 5000, labels = data_UK$Freq, pos = 3, cex = 0.8, col = "black", xpd = TRUE)
```

#### Road type crashes

```{r}
data_UK <- data.frame(table(tidy_num$Road_type))
data_UK <- data_UK[order(-data_UK$Freq), ]

# Increase the bottom and right margins
par(mar=c(8.2, 4, 3, 3))  # Adjust the second and fourth values as needed

barplot(data_UK$Freq, names.arg = data_UK$Var1, 
        las = 2, col = "skyblue", 
        main = "Accident Counts by Road type", 
        xlab = "", 
        ylab = "",
        ylim = c(0, 700000))

# Add the number of accidents above each column with adjusted x-coordinates
text(x = seq_along(data_UK$Var1)*1.22-0.55, y = data_UK$Freq + 5000, labels = data_UK$Freq, pos = 3, cex = 0.8, col = "black", xpd = TRUE)
```

### Austin

#### Traffic Control Type crashes

```{r}
data_austin <- data.frame(table(new_df$Traffic.Control.Type))
data_austin <- data_austin[order(-data_austin$Freq), ]
par(mar=c(15, 3, 3, 1)) # 15 line height for bottom margin
barplot(data_austin$Freq, names.arg = data_austin$Var1, 
        las = 2, col = "skyblue", 
        main = "Accident Counts by Traffic Control Type", 
        xlab = "", 
        ylab = "Frequency")
text(x = seq_along(data$Var1)*1.22-0.55, y = data_austin$Freq, labels = data_austin$Freq, pos = 3, cex = 0.8, col = "black", xpd = TRUE)
```

#### Daily crash frequency

```{r}
data_austin <- data.frame(table(new_df$Day.of.Week))
data_austin <- data_austin[order(-data_austin$Freq), ]
par(mar=c(6, 3, 3, 1)) # 15 line height for bottom margin
barplot(data_austin$Freq, names.arg = data_austin$Var1, 
        las = 2, col = "skyblue", 
        main = "Accident Counts by Day of the week", 
        xlab = "", 
        ylab = "Frequency")
text(x = seq_along(data_austin$Var1)*1.22-0.55, y = data_austin$Freq, labels = data_austin$Freq, pos = 3, cex = 0.8, col = "black", xpd = TRUE)
```

### Motorbikes

#### Age to crash density

## Plotting

### Plotting function

```{r}
generate_mean_plot_beatiful <- function(data, x_col, y_col, title, x_label, y_label) {
    mean_df <- aggregate(data[[y_col]] ~ data[[x_col]], data = data, FUN = mean)
    
    ggplot(mean_df, aes(x = .data[["data[[x_col]]"]], y = .data[["data[[y_col]]"]])) +
      geom_line(color = "blue", size = 1) +
      geom_ribbon(aes(ymin = min(.data[["data[[y_col]]"]])-0.01, ymax = .data[["data[[y_col]]"]]), fill = "blue", 
                  alpha = 0.3) + 
      labs(x = x_label, y = y_label, title = title) +
      theme_minimal() +
      theme(axis.text = element_text(size = 12),
            axis.title = element_text(size = 14, face = "bold"),
            plot.title = element_text(size = 16, face = "bold")) +
      theme(panel.grid.major = element_line(color = "gray", linetype = "dashed"),
            panel.grid.minor = element_blank())

}

generate_mean_plot <- function(data, x_col, y_col, title, x_label, y_label) {
    mean_df <- aggregate(data[[y_col]] ~ data[[x_col]], data = data, FUN = mean)
    
    ggplot(mean_df, aes(x = .data[["data[[x_col]]"]], y = .data[["data[[y_col]]"]])) +
        geom_point(position = position_dodge(width = 0.2)) +
        labs(x = x_label, y = y_label, title = title)
    
}
```

### UK

#### Injuries vs speed limits

```{r}
generate_mean_plot_beatiful(selected_col_num, "Speed_limit", "Severity_num",
                   "Mean injury at different speed limits",
                   "Speed Limit",
                   "Mean injury")
```

#### Severity vs light situation

```{r}
generate_mean_plot_beatiful(selected_col_num, "Light_conditions_num", "Severity_num",
                   "Mean injury in different light condition",
                   "Light conditions",
                   "Mean injury")
```

Mean light condition vs severity

```{r}
generate_mean_plot_beatiful(selected_col_num, "Severity_num", "Light_conditions_num",
                   "Mean light conditions for different injuries",
                   "Mean injury",
                   "Light conditions")
```

### Austin

##### Injuries vs speed limits

```{r}
# Use speed limit as x, and y as Severity
generate_mean_plot(new_df, 
                    "Speed.Limit", "Crash.Numeric",
                   "Mean of overall crash Severity for Speed Limit",
                   "Speed Limit",
                   "Mean Crash Severity")
```

#### Hypothesys

```{r}
df_severity_speed <- aggregate(new_df[["Crash.Numeric"]] ~ new_df[["Speed.Limit"]], data = new_df, FUN = mean)
colnames(df_severity_speed) <- c('Speed.Limit','CrashMean') 

# Attempt to build linear regression
relation <- lm(df_severity_speed$CrashMean~df_severity_speed$Speed.Limit)
y <- df_severity_speed$CrashMean
x <- df_severity_speed$Speed.Limit
ggplot(df_severity_speed, aes(Speed.Limit, CrashMean)) +
        geom_point(position = position_dodge(width = 0.2)) +
        labs(x = "Speed Limit", y = "Crash Severity Mean", title = "Speed Limit & Crash Mean(Linear Regression)") + 
        geom_smooth(method = "lm", se = FALSE)
       

# Attemplt to build polyniomial regression
df_severity_speed$Speed.Limit2 = df_severity_speed$Speed.Limit^2

poly_reg = lm(formula = CrashMean ~ .,data = df_severity_speed)

x_grid = seq(min(df_severity_speed$Speed.Limit), max(df_severity_speed$Speed.Limit), 0.1)
ggplot() +
  geom_point(aes(x = df_severity_speed$Speed.Limit, y = df_severity_speed$CrashMean),
             colour = 'black') +
  geom_line(aes(x = x_grid, y = predict(poly_reg,
                                        newdata = data.frame(Speed.Limit = x_grid,
                                                             Speed.Limit2 = x_grid^2))),
            colour = 'blue') +
  ggtitle('Speed Limit & Crash Mean (Polynomial Regression)') +
  xlab('Speed Limit') +
  ylab('Crash Severity Mean')
```

##### Injuries vs light situation

```{r}
new_df$Light.Numeric <- ifelse(new_df$Light.Condition=="Dusk", 5,
                    ifelse(new_df$Light.Condition=="Daylight", 4,
                    ifelse(new_df$Light.Condition=="Dawn", 3, 
                    ifelse(new_df$Light.Condition=="Dark, Not Lighted", 2,
                    ifelse(new_df$Light.Condition=="Dark, Lighted", 1, 6)))))
# Delete unknown values
new_df <- subset(new_df, Light.Numeric != 6)
generate_mean_plot(new_df, 
                    "Crash.Time", "Light.Numeric",
                   "Time of the day and light comdition",
                   "Time of the day",
                   "Light quality higher -> better")
```

### Motorbikes

##### Age to crash density

```{r}
pie_diagram = function(list)
{
  vals = c()
  names = c()
  for (i in unique(list))
  {
    vals = append(vals, length(list[list==i]))
    names = append(names, i)
  }
  pie(vals,names)
}
hist_diagram = function(list)
{
  hist(x=strtoi(list), breaks=100)
}
plot_all = function(data)
{
  pie_names = c("AmbulanceR", "BikeInjury","RuralUrban","BikeSex",       "TraffCntrl",  "SpeedLimit",
  "LightCond",  "Region" ,    "Weather" )
  hist_names = c("BikeAge", "CrashHour")
  for (i in pie_names)
    pie_diagram(as.list(data[i])[[1]])
  for (i in hist_names)
    hist_diagram(as.list(data[i])[[1]])
}
sum_of_normal_d = function(data, label = "")
{
  list = as.list(data["BikeAge"])[[1]]
  N=10000
  new_data = rnorm(N, mean = c(18,49), sd=c(7,11))
  hist(x=strtoi(list), prob=TRUE,breaks=40, main = label, xlab = "Age")
  lines(density(new_data), col = "green", lwd = 4)
}

sum_of_normal_d(data, "Data of the all time")
```

#### Hypotesys

```{r}
for (i in 2007:2019)
  sum_of_normal_d(data[data$CrashYear==i,], paste("Data of the year ", as.character(i)))
```
